Parallel computations for the multidimensional multiextremal optimization problems with the dimensionality reduction using Peano curves

### 2017-10-03 2017/ParOptBook/Chapter6.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Software Systems for Global Optimization }
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Name of First Author and Name of Second Author}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\section{Section Heading}
Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.

>>>>

%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%\usepackage{amsmath}

%\usepackage[colorlinks=true]{hyperref}
%\hypersetup{urlcolor=blue, citecolor=red}

%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Parallel computations for the multidimensional multiextremal optimization problems with the dimensionality reduction using Peano curves }
\titlerunning{Parallel computations for the multidimensional optimization}

\author{Konstantin Barkalov, Victor Gergel, Roman Strongin}
\authorrunning{K. Barkalov, V.Gergel, R. Strongin}

\institute{Konstantin Barkalov,  Victor Gergel, Roman Strongin \at Lobachevsky State University of Nizhni Novgorod,  Nizhni Novgorod, Russia \email{konstantin.barkalov@itmm.unn.ru}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{The chapter considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents). The initial multidimensional optimization problem is substituted by an equivalent one-dimensional one based on the application of the mapping of a multidimensional search domain onto a segment of the real axis. A general description of the approach and its substantiations are given. The development of the classical scheme based on the application of a set of mappings and the implementation of new types of evolvents is considered for the parallel algorithm.}

\abstract{The chapter considers dimensionality reduction scheme based on Peano-type space-filling cures (evolvents). The initial multidimensional optimization problem is substituted by an equivalent one-dimensional one based on the application of the mapping of a multidimensional search domain onto a segment of the real axis. A general description of the approach and its substantiations are given. The development of the classical scheme based on the application of a set of mappings and the implementation of new types of evolvents is considered for the parallel algorithm.}

\section{General approach scheme}

Let us consider a multidimensional global optimization problem
\begin{equation}\label{6_problem} 
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_j(y)\leq 0, \; 1 \leq j \leq m\right\}},
\end{equation} 
\begin{equation}\label{6_D}
D=\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1\leq i \leq N \right\}.
\end{equation}
This problem statement covers a large class of problems since any hyperinterval 
\[
S=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}
\]
can be reduced to the  hypercube (\ref{6_D}) by linear transformation of coordinates.

There is a number of ways to adapt efficient one-dimensional algorithms for solving multidimensional problems; see, for example, the diagonal partitions method in \cite{6_Sergeyev2006,6_Sergeyev2015,6_Sergeyev2017} or the simplicial partitions method in \cite{6_Zilinskas2008,6_Zilinskas2014,6_Zilinskas2014_1}.

The dimensionality reduction method described below uses a mapping of a multidimensional search domain onto a one-dimensional interval by means of so called \textit{Peano curves}. Peano, an Italian mathematician, has found that an single-valued continuous mapping $y(x)=(y_1(x),y_2(x))$ of the interval $[0,1]$ onto the square 
\[
\left\{y\in R^2: -2^{-1}\leq y_1,y_2 \leq 2^{-1} \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\}
\]
can be constructed.
This result was generalized onto a multidimensional case, i.e., the existence of the curves $y(x)$ defined by continuous coordinate functions $y_i(x), \; x \in [0,1],\; 1 \leq i \leq N,$ and mapping uniquely the unit interval $[0,1]$ onto a $N$-dimensional hypercube 
\[
D=\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1\leq i \leq N \right\} = \left\{ y(x): 0 \leq x \leq 1 \right\}
\]
had been proven. These curves, called also \textit{Peano curves}, allow to reduce a multidimensional constrained optimization problem over the domain $D$ to a one-dimensional constrained minimization problem over the unit interval $[0,1]$
\begin{equation}\label{6_problem_1} 
\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)):x\in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}}.
\end{equation}

The considered dimensionality reduction scheme juxtaposes to a multidimensional problem with Lipschitz objective function and Lipschitz constraints a one-dimensional problem, where the corresponding functions satisfy uniform H\"{o}lder condition (see \cite{6_Strongin2000,6_Strongin2013}), i.e.,
\begin{equation}\label{6_Holder}
\left|g_j(y(x'))-g_j(y(x''))\right| \leq K_j \left|x'-x''\right|^{1/N}, \; x',x''\in [0,1], \; 1\leq j \leq m+1. 
\end{equation}
Here $N$ is the dimensionality of the initial multidimensional problem and the coefficients $K_j$ are related with Lipschitz constant $L_j$ of the initial problem as 
\[
K_j \leq 2L_j \sqrt{N+3},\; 1\leq j \leq m+1.
\]
Some issues of H\"older functions optimization are considered in \cite{6_Gourdin,6_Lera2002,6_Lera2010,6_Hime,6_Lera2015}.

We will consider initial problem (\ref{6_problem}) assuming that the problem functions $g_j(y),\; 1 \leq j \leq m+1,$ may be defined partially. This means that they are defined and computable only in the subranges $Q_j \in [0,1]$, 
\begin{equation}\label{6_Q}
Q_1=[0,1], \; Q_{j+1}=\left\{x \in Q_j : g_j(y(x)) \leq 0 \right\}, \; 1 \leq j \leq m.
\end{equation}
These conditions allows to introduce a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints computed at this point. The index $\nu(x)$ can also be defined by the conditions
\begin{equation}\label{6_nu}
g_j(y(x)) \leq 0, \; 1 \leq j < \nu, \; g_\nu(y(x))>0,
\end{equation}
where the last inequality is inessential if $\nu=m+1$.

Therefore, the execution of an iteration at the point $x \in [0,1]$ is reduced to the successive computation of the values $g_j(y(x)),\; 1 \leq j \leq \nu(x),$ i.e., the next value $g_{j+1}(y(x))$ is computed if and only if $g_j(y(x)) \leq 0$. The process of computations is terminated either as a result of occurrence of the inequality $g_j(y(x))>0$ or by satisfying the equality $\nu(x)=m+1$. Moreover, the described procedure called the trial at the point $x \in [0,1]$ leads automatically to determining the index $\nu$ of this point. %since conditions (\ref{6_nu}) are equivalent to the inequalities 
The pair of values 
\[
\nu=\nu(x), \; z=g_\nu(y(x)),
\]
generated by the trial at the point $x \in [0,1]$, is called \textit{the trial result}.

Besides the exact solution $y^\ast$ from (\ref{6_problem}) let us consider also the \textit{$\epsilon$-reserved solution} of the problem (\ref{6_problem}) defined by the conditions 
\begin{equation}\label{6_eps_r}
\varphi(y_\epsilon) = \min \left\{ \varphi(y): \; y \in D, \; g_j(y) \leq -\epsilon_j, \; 1 \leq j \leq m \right\},
\end{equation}
where $\epsilon_1,...,\epsilon_m$ are the positive values (the ``reserves'' with respect to each constraint). Let us introduce into the consideration also the set
\begin{equation}\label{6_eps_solution}
Y_\epsilon =  \left\{ y \in D: \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_\epsilon) \right\}
\end{equation}
of all feasible points of problem (\ref{6_problem}), which are not worse (with respect to the objective function value) than the $\epsilon$-reserved solution. Fig.~\ref{6_fig_1} illustrates the definition introduced.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.7\linewidth]{figures/6_1.png}
\caption{$\epsilon$-reserved solution}
\label{6_fig_1}     
\end{figure}

The existence of the $\epsilon$-reserved solution of problem (\ref{6_problem}) can be interpreted as an analog of the regularity conditions in the classical problems of nonlinear programming. It should be noted also the applied importance of this condition. Even if the exact solution $y^\ast$ is known, its practical realization is possible as some approximation of $y^\ast$ only. Therefore, it is important that there exist the points from the feasible domain $Q$ close to $y^\ast$ (with respect to the coordinates and values). The existence of the $\epsilon$-reserved solution guarantees the existence of such points. In this case, the domain $Y_\epsilon$ can play a role of the set of suitable approximations.

Prior to going to the detailed description of the Peano curves and of the algorithm, let us consider some properties of reduced problem (\ref{6_problem_1}), which are absent in the pure one-dimensional case. First, as it has been already mentioned above, the functions $g_j(y(x)), \; 1 \leq j \leq m+1,$ are not Lipschitzian. They satisfy H\"older condition. Consequently, all the interval lengths in the search algorithm rules have to be replaced by the lengths in the metric
\[
\rho (x',x'') = \left(\left|x'-x''\right|\right)^{1/N}, \; x',x'' \in [0,1].
\]
Second, in general, sets (\ref{6_Q}) being the inverse images of the essentially non-cubic $N$-dimensional subsets with the nonlinear boundaries $g_j(y)=0, \; y \in D, \; 1\leq j \leq m,$ are not the finite unions of $x$-intercepts. Consequently, taking into account the $\epsilon$-reserved solutions plays a significantly more important role for the convergence in the reduced univariate problem (\ref{6_problem_1}) than in the pure one-dimensional case.

The issues of the numerical construction of the Peano-type space filling curves and the corresponding theory are considered in details in \cite{6_Butz,6_Sagan,6_Strongin2000,6_Strongin2013}. Here we will describe the basic steps of constructing a curve mapping a unit interval of the real axis $[0,1]$ onto a hypercube  
\[
D=\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1\leq i \leq N \right\}.
\]

\begin{enumerate}
	\item 
A hypercube $D$ with the unit edge length is divided by the coordinate hyperplanes into $2^N$ hypercubes of the first partition (with the edge length equal to $1/2$), which are indexed by the numbers $z_1$ ranging from $0$ to $2^N-1$. Let us agree to designate the hypercube of the first partition with the index $z_1$ as $D(z_1)$.

Next, each hypercube of the first partition is also divided into $2^N$ hypercubes of the second partition (with the edge length equal to $1/4$) by the hyperplanes parallel to the coordinate axes and crossing the middles of the hypercube edges orthogonal to these hyperplanes. The hypercubes of the second partition  included in the hypercube $D(z_1)$ are indexed by the numbers $z_2$ ranging from $0$ to $2^N-1$. The hypercube of the second partition  with the index $z_2$ from  $D(z_1)$ is denoted as $D(z_1, z_2)$. The case of $N=2$ is presented in Fig.~\ref{6_fig_2} for $m=1$, $m=2$.

\begin{figure}
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=0.9\linewidth]{figures/6_2a.png} \\ $m=1$}
\end{minipage}
\hfill
\begin{minipage}{0.5\linewidth}
\center{\includegraphics[width=0.9\linewidth]{figures/6_2b.png} \\ $m=2$}
\end{minipage}
\caption{The partition of an $N$-dimensional hypercube}
\label{6_fig_2}
\end{figure}

Continuing this process, one can construct the hypercubes of any $m$-th partition with the edge length equal to $(1/2)^m$, which are denoted as $D(z_1,...,z_m)$. It is obvious, that $D(z_1) \supset D(z_1,z_2) \supset ... \supset D(z_1,...,z_m)$ and $0 \leq z_j \leq 2^N-1$, \mbox{$1 \leq j \leq m$}.

\item
Now let us perform the division of the interval $[0,1]$ into $2^N$ equal parts, each of them also is divided into $2^N$ equal parts, etc. The elements of each partition are indexed from the left to the right by the numbers $z_j$ ranging from $0$ to \mbox{$2^N-1$} where $j$ is the index of subdivision. Let us denote the intervals of the $m$-th partition as $d(z_1,...,z_m)$ where, for example, $d(z_1,z_2)$ means the interval of the second partition with the index $z_2$ being a part of the interval $d(z_1)$ of the first partition with the index $z_1$.

One can note that  $d(z_1) \supset d(z_1,z_2) \supset ... \supset d(z_1,...,z_m)$ and the length of the interval $d(z_1,...,z_m)$ is equal to  $(1/2)^{mN}$. An interval $d(z_1,...,z_m)$ is supposed to include its left hand end. It contains its right hand boundary if and only if when $z_1=z_2=...=z_m=2^N-1$. The case of $N=2$ is presented in Fig.~\ref{6_fig_3} for $m=1$ and $m=2$.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.9\linewidth]{figures/6_3.png}
\caption{Partition of a one-dimensional interval}
\label{6_fig_3}     
\end{figure}

\item
Assume that the point $y(x) \in D$ corresponding to the point $x \in [0,1]$ for any $m \geq $ belongs to the hypercube $D(z_1,...,z_m)$ if $x$ falls into the interval $d(z_1,...,z_m)$, i.e.,
\[
x \in d(z_1,...,z_m) \rightarrow y(x) \in D(z_1,...,z_m).
\]
The constructed correspondence  $y(x)$ is single-valued.
\item
In order to provide the continuity of the constructed correspondence, let us impose the following requirements on the indexing order of the hypercubes of each partition. Since $2^{mN}$ centers $y(z_1,...,z_m)$ of the hypercubes of the $m$-th partition $D(z_1,...,z_m)$ form a uniform orthogonal grid in the domain $D$ (the step of this grid with respect to any coordinate is equal to $2^{-m}$), one can introduce the following numeration of the grid nodes. Let us enumerate from the left to the right by index $i$ all subintervals of the $m$- 	th partition in the interval $[0,1]$, i.e.,
\[
d(z_1,...,z_m)=[x_i, x_{i+1}), \; 0 \leq i < 2^{mN}-1,
\]
were the left end of the $i$-th interval is denoted as $x_i$.

Assume the center of the hypercube $D(z_1,...,z_m)$ to have the same index $i$ as the interval $d(z_1,...,z_m)$ corresponding to this hypercube, i.e.,
\[
y_i = y(z_1,...,z_m), \; 0 \leq i < 2^{mN}-1.
\]
The centers $y_i$ and $y_{i+1}$ correspond to the adjacent hypercubes having a common facet. More details on the indexing of the hypercubes can be found in \cite{6_Strongin2013}.
\end{enumerate}

Let us consider the mapping  $l(x)$ of the interval $[0,1]$ into a hypercube $D$ defined by the expression 
\[
l(x)=y_i+(y_{i+1}-y_i)\frac{w(x)-x_i}{x_{i+1}-x_i}, \; x_i \leq w(x) \leq x_{i+1},
\]
\[
w(x)=x(1-2^{-mN}), \; 0 \leq x \leq 1.
\]
The image of any subinterval
\[
\left[x_i(1-2^{-mN})^{-1}, x_{i+1}(1-2^{-mN})^{-1}\right], \; 0 \leq i < 2^{mN}-1
\]
of the interval $[0,1]$ for the correspondence $l(x)$ is a linear segment connecting the nodes $y_i$ and $y_{i+1}$.

Thus, we have built a piecewise linear curve $l_m(x), 0 \leq x \leq 1,$ connecting the nodes $y_i, 0 \leq i< 2^{mN}-1,$ in the order of their indexing. This curve obtained numerically (\textit{the evolvent})  is an approximation of the theoretical Peano curve with the precision not worse than $2^{-m}$ with the respect to each coordinate (the parameter m is called\textit{ the density of the evolvent}). As an illustration, an image of the interval $[0,1]$ for the mapping $l_m(x)$ in the case $N=2, m=3$ is presented in Fig.~\ref{6_fig_4}. The grid nodes are marked with the dark circles.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.5\linewidth]{figures/6_4.png}
\caption{A piecewise linear curve}
\label{6_fig_4}     
\end{figure}

\section{Multidimensional generalized global search algorithm}
The considered dimensionality reduction scheme associates a multidimensional problem with Lipschitz objective function and Lipschitz constraints with a one-dimensional problem, where the corresponding functions satisfy uniform H\"older condition (\ref{6_Holder}). As result, all lengths of the intervals appearing in the rules of the search algorithm should be replaced by the lengths in a new metrics, where the distance is defined by the expression
\[
\Delta_i = \left|x_i-x_{i-1}\right|^{1/N}.
\]
Also, in the reduction of the problem dimensionality by Peano curve  $y(x), 0 \leq x \leq 1,$ a trial at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist in the following sequence of operations.
\begin{itemize}
	\item Determine the image  $y^k=y(x^k)$ in accordance with the mapping  $y(x)$.
	\item Compute the values  $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ are determined by the conditions
	\[
	g_j(y^k)\leq 0, \; 1 \leq j<\nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
	The occurrence of the first violation of the constraint terminates the trial at the point $y^k$. In the case, when the point $y^k$ is a feasible one, i.e., when $y(x^k) \in Q_{m+1}$, the trial includes the computation of the values of all functions of the problems and the index is accepted to be $\nu = m+1$. The pair of values
	\begin{equation}\label{6_trial_res}
	\nu = \nu(x^k), \; z^k = g_\nu(y(x^k))
	\end{equation}
	is a \textit{result of the trial} at the point $x^k$.
\end{itemize}

Thus, one can easily modify the initial one-dimensional algorithm and obtain on this base a multidimensional \textit{generalized index global search algorithm}, described below.

The first trial is executed at an arbitrary internal point $x^1 \in (0,1)$. The selection of the point $x^{k+1}, k \geq 1,$ of any next trial is determined by the following rules.

Rule 1. Renumber the points $x^1,...,x^k$ of the preceding trials by the
lower indices in ascending order of coordinate values, i.e.
\begin{equation}\label{6_trial_points}
0=x_0<x_1<\dots <x_k<x_{k+1}=1,
\end{equation}
and juxtapose to them the values $z_i=g_\nu(y(x_i)), \; \nu=\nu(x_i), \; 1 \leq i \leq k$, from (\ref{6_trial_res}) computed at these points. The points $x_0=0$ and
$x_{k+1}=1$ are introduced additionally for the convenience of further notations, while the values $z_0$ and
$z_{k+1}$ are not defined.

Rule 2. Classify the indices $i, \; 1 \leq i \leq k$, of the trial points (\ref{6_trial_points})
according to the number of the problem constraints fulfilled at these points by constructing the sets
\[
I_\nu =\left\{i:1 \leq i \leq k, \; \nu=\nu(x_i) \right\}, \; 1 \leq \nu \leq m+1,
\]
containing the numbers of all the points $x_i, \; 1 \leq i \leq k$, with
the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are
interpreted as the ones having indices equal to zero. An additional set $I_0=\left\{0,k+1\right\}$ corresponds to them.

Determine the maximum value of the index
\[
M=\max\left\{\nu(x_i), \; 1 \leq i \leq k \right \}.
\]

Rule 3. Compute the current lower estimates
\begin{equation}\label{6_mu}
\mu = \max\left\{ \frac{\left|z_i-z_j\right|}{ (x_i - x_j)^{1/N} }, \; i,j \in I_\nu, \; i>j \right\}
\end{equation}
for the unknown H{\"o}lder constants $K_\nu$ of the functions $g_\nu(y),1
\leq \nu \leq m+1$. If a set $I_\nu$ contains less than two elements, or
if $\mu_\nu$ from (\ref{6_mu}) is equal to zero, then assume $\mu_\nu=1$.

Rule 4. For all nonempty sets $I_\nu, \; 1 \leq \nu \leq m+1$, compute the
estimates
\[
z_\nu^\ast = \left\{
   \begin{array}{lr}
     -\epsilon_\nu, & \nu < M,\\
     \min\{ g_\nu(y(x_i)): i\in I_\nu \}, & \nu = M,
   \end{array}
\right.
 \]
where the vector $\epsilon_R = (\epsilon_1,...,\epsilon_m)$ is a predefined vector with the nonnegative components (\textit{the reserve vector}).

Rule 5. For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute
the \textit{characteristics} $R(i)$ :
\[
R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
\]
\begin{equation}%\label{eq:14}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
\end{equation}
\[
R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
\]
where $\Delta_i=(x_i - x_{i-1})^{1/N}$. The values $r_\nu > 1, \; 1 \leq
\nu \leq m+1,$ are parameters of the algorithm. An appropriate selection
of $r_\nu$ allows to consider the product $r_\nu \mu_\nu$ as an estimate
of the H{\"o}lder constants $K_\nu, \; 1 \leq \nu \leq m+1$.

Rule 6. Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}\label{6_MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}

Rule 7. Make the next trial at the midpoint of the interval
$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not
the same, i.e.,
\[
x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\]
Otherwise, make the trial at the point
\begin{equation}%\label{eq:142}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}

We may take as termination condition the inequality $\Delta_t \leq
\epsilon$, where $t$ is from (\ref{6_MaxR}) and $\epsilon>0$ is the
predefined accuracy.

\textbf{The algorithm convergence conditions.} These conditions are a special case of the convergence theorem which will be proved in Subsection 6.3 for the parallel algorithm with a single curve on a single processor.

\begin{example} \label{6_example1}
As an illustration, let us consider the problem of minimization of the function
\begin{align*}
 \varphi(y_1,y_2) = &-1.5y_1^2\exp(1-y_1^2-20.25(y_1-y_2)^2)- \\
				            &-(0.5(y_1-1)(y_2-1))^4\exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
\end{align*}
within the domain $0 \leq y_1\leq 4,\ -1\leq y_2\leq 3$, with the constraints 
\begin{align*}
g_1(y_1,y_2)&=0.01((y_1-2.2)^2+(y_2-1.2)^2-2.25)\leq 0,\\
g_2(y_1,y_2)&=100(1-(y_1-2)^2/1.44-(0.5y_2)^2)\leq 0,\\
g_3(y_1,y_2)&=10(y_2-1.5-1.5\sin(6.283(y_1-1.75)))\leq 0.
\end{align*}

Fig.~\ref{6_fig_5} shows the square search domain and the points of $1098$ trials obtained by the index method with the following parameters: the reliability parameters $r_1=...=r_4=2$, the accuracy $\epsilon = 10^{-3}$, the evolvent density $m=12$.
\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.8\linewidth]{figures/6_5.png}
\caption{The results of solving the test problem}
\label{6_fig_5}     
\end{figure}

\end{example}

Note that in applied problems, estimating the values of the constraints and of the objective functions very often requires considerable computation resources. The results of the experiments confirm the efficiency of the index scheme of the accounting for the constraints since the values of the functions $g_1,\ g_2,\ g_3,\ \varphi = g_4$ were computed $k_1=1098,\ k_2=623,\ k_3=392,$ and $k_4=152$ times, respectively. If the search of the solution were conducted on a uniform grid, in order to achieve the same accuracy $\epsilon=10^{-3}$ $k_1=1.6\cdot 10^7,\ k_2=7 \cdot 10^6,\ k_3=3\cdot 10^6,$ and $k_4=1.4 \cdot 10^6$ computations of the values of the functions $g_1,\ g_2,\ g_3,\ \varphi = g_4$, respectively, would be necessary.

\section{Parallel multidimensional multiextremal methods based on multiple Peano curves}

\subsection{Application of multiple mappings}\label{6_section_shift}

The reduction of the multidimensional problems to the one-dimensional ones using the Peano curves has such important properties as the continuity and the uniform boundedness of the function differences for limited variation of argument. However, a partial loss of information on the nearness of the points in the multidimensional space takes place since a point $x \in [0,1]$ has the left and the right neighbors only while the corresponding point $y(x) \in D \subset R^N$ has the neighbors in $2N$ directions. As a result, when using the mappings like Peano curve the images $y',\ y'',$ which are close to each other in the $N$-dimensional space can correspond to the preimages $x',\ x'',$ which can be far away from each other in the interval $[0,1]$. This property results in the excess computations since several limit points $x',\ x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.

One of the possible ways to overcome this disadvantage consists in using the multiple mappings
\begin{equation}%\label{eq:142}
Y_L(x)=\left\{y^0(x),\ y^1(x),...,\ y^L(x)\right\}
\end{equation}
instead of single Peano curve $y(x)$ (see \cite{6_Strongin1991,6_Strongin1992,6_Strongin2000}).

Let us consider a family of the hypercubes 
\begin{equation}\label{6_hypercubes}
D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L,
\end{equation}
where the hypercube $D_{l+1}$ is obtained by the translation of the hypercube $D_l$ along the main diagonal with displacement equal to $2^{-l}$ in each coordinate. In Fig.~\ref{6_fig_6} the hypercubes $D_0,...,D_3$ for the case $N=2,\ L=3$ are presented.

\begin{figure}[t]
%\sidecaption[t]
\includegraphics[width=0.7\linewidth]{figures/6_6.png}
\caption{Multiple mappings}
\label{6_fig_6}     
\end{figure}

Let us assume that the evolvent $y^0(x)$  maps the interval $[0,1]$ onto the hypercube $D_0$ from (\ref{6_hypercubes}), i.e.,
\[
D_0 = \left\{y^0(x) : x \in [0,1]\right\}.
\]
Then, the evolvents $y^l(x)=\left\{y_1^l(x),...,y_N^l(x)\right\}$, the coordinates of which are defined by the conditions 
\[
y_i^l(x)=y_i^{l-1}(x)+2^{-l},\ 1\leq i\leq N, \ 1\leq l\leq L,
\]
map the interval $[0,1]$ onto the corresponding hypercubes $D_l,\ 1\leq l \leq L$. In Fig.~\ref{6_fig_6} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\ x\in [0,1],$ is shown as the broken line. Since the hypercube $D$ from (\ref{6_D}) is included in the common part of the family of hypercubes (\ref{6_hypercubes}) (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{6_fig_6}), having introduced an additional constraint function
\begin{equation}\label{6_g0}
g_0(y)=\max\left\{\left|y_i\right| - 2^{-1}:\ 1\leq i\leq N\right\},
\end{equation}
one can present the initial hypercube $D$ in the form
\[
D=\left\{y^l(x):\; x\in [0,1],\ g_0(y^l(x))\leq 0 \right\},\ 0\leq l \leq L,
\]
i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$ has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.

Thus, each evolvent $y^l(x),\ 0\leq l \leq L,$ generates its own problem of the type (\ref{6_problem_1}) featured by its own extended (in comparison with $D$) search domain $D_l$ and the additional constraint with the left hand part from (\ref{6_g0})
\begin{equation}\label{6_problem_l} 
\min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 0 \leq j \leq m\right\}}, \ 0 \leq l \leq L.
\end{equation}

The problems (\ref{6_problem_l}) correspond to the domains $Q_0^l=[0,1]$ and $Q_{j+1}^l, 0 \leq j \leq m,$ defined by the expression 
\[
Q_{j+1}^l = \left\{x \in Q_j^l:g_j(y^l(x))\leq 0\right\},\ 0\leq j\leq m,
\]
for the corresponding evolvents $y^l(x)$. The application of a multiple mapping defines the following relation for the nearness in the multidimensional search domain and in the one-dimensional one.

\begin{theorem}
Let a point $y^\ast$ from the domain $D$ be contained in the line segment with the end-points $y',y'' \in D$ meeting the requirements 
\[
\left|y'_j - y''_j\right|\leq 2^{-p},\ y'_i=y''_i=y^\ast_i, \ 1\leq i \leq N,\ i \neq j,
\]
where $p$ is an integer  and $1\leq p \leq L$, i.e., the line segment is collinear with the $j$-th axis in $R^N$. Then, there exist at least one mapping $y^l(x),\ 0\leq l\leq L$, and the preimages $x^\ast,\; x',\; x''\in [0,1]$ such that
\[
y^\ast = y^l(x^\ast),\ y'=y^l(x'),\ y'' = y^l(x'')
\]
and
\[
\max \left\{ \left|x'-x^\ast\right|,\; \left|x''-x^\ast\right|,\; \left|x'-x''\right|\right\} \leq 2^{-pN}.
\]
\end{theorem}
\begin{proof}
Proof of this theorem is given in \cite{6_Strongin2000}.
\qed
\end{proof}
